Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2021-01-10 papers/thesis-kirstie/thesis.tex
edits

%Newer Intro....
%The goal of this project is to generate temperature-density 
%and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
%(WCA) model fluid using classical Density Functional Theory (cDFT). 
%To do this, the Helmholtz free energies of the fluid
%in a liquid state and in a crystalline solid state must be found.
%In general, the Helmholtz free energy, which consists of an ideal part 
%and an excess part, can be written as a functional of the number density 
%profile $n(\vec{r})$ which describes the way the average number 
%density varies spatially. 
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

@@

A liquid will eventually freeze as heat is removed, or the liquid is compressed. 
The state of a simple fluid can be characterized by its entropy per particle~\cite{entropyfingerprint}.
The entropy of a pure substance in a crystalline state is higher than the entropy 
that it would have in its liquid state at a given temperature and pressure, 
since there is more freedom for a particle to move between atoms  
at lattice sites than in a dense liquid with irregular and blocked pathways~\cite{Hansen}.
Both the entropy and the volume (and thus the density) change abruptly 
at the phase transition.
The particular structure and density that result are affected 
by the interatomic potential. When repulsive, soft sphere fluids 
(that are not too soft) freeze, they form lattice structures composed of 
Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
This structure is what one would see by filling a box with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

\begin{figure}
    \centering
    \includegraphics[height=8cm]{figs/P-T_Diagram.pdf}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
\end{figure}

\textcolor{red}{A phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure~\ref{fig:P-T_Diagram}. 
%
Points along the line that separates the solid and the liquid phases 
indicate a phase transition where the solid and liquid phases coexist, 
and show that both the solid and the liquid have the same temperature (necessary 
for thermal equilibrium)
and the same pressure (necessary for mechanical 
equilibrium)at a phase transition.
%
, and the same chemical potential (necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some arbitrary, unknown substance composed of atoms, 
or molecules which exhibit both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will just simply
be a fluid, which we will refer to as a liquid in this paper.}
%which may be referred to as a "liquid", since at the phase boundary the fluid is dense.

@@

 \begin{figure}
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
    \caption{Liquid and solid densities at the solid-liquid phase 
    transition are identified using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. Each curve consists of a 
collection of data points $(\frac{F}{N}, \frac{1}{n})$ for the liquid and
the solid respectively, which will be obtained by using cDFT to find the 
Helmholtz free energies of the liquid and the solid for various 
temperatures and number densities. 
Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transition, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the points at which the 
tangents to each curve have the same slope, as shown by the black line. 
This process is called Maxwell's Double Tangent Construction.

To facilitate finding the pressure at the phase transition computationally, 
it is helpful to consider the Gibbs free energy given by G=F+PV. 
Rearranging the terms and dividing by N gives an  expression for the 
Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
per atom  $g$ (which is the chemical potential)  
\begin{equation}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{equation}
\begin{align}f=-P\frac{1}{n}+g\end{align}
This equation has the form of a line
\begin{equation}
    y=mx+y_o
\end{equation}
and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom, or chemical potential given by the y-intercept. 
At the point of phase transition the chemical potential of the liquid is 
the same as the chemical potential of the solid, as is the pressure.
\begin{equation}
    \mu_L=\mu_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}
\end{equation}

 \begin{figure}
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs 
    when $\mu_L=\mu_S$.}
    \label{fig:GibbsvsP}
  \end{figure}

\begin{figure}
    \centering
    \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
    \caption{The liquid and solid densities at the phase transition are 
    found from a plot of pressure vs inverse density for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

The pressure at the phase transition can be obtained from the point of 
intersection that occurs when the chemical potential 
is plotted against the pressure for both the liquid and the 
solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
This plot is constructed from data collected from lines tangent 
along the curves on the Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
liquid and the solid at a given temperature. 
The slope of the tangent line at a given point gives the negative of the 
pressure at that point, and the corresponding y-intercept of the tangent 
line gives the chemical potential at that point. Collecting this data 
(P,~$\mu$) at each point for the liquid and the solid respectively, 
produces the curves in Figure~\ref{fig:GibbsvsP}.

\begin{figure}
    \centering
    \includegraphics[height=8cm]{figs/T-n_Diagram.pdf}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the liquid and the solid during a phase 
    transition. The gray region is where different phases coexist in 
    varying amounts at their phase transition number densities.}
    \label{fig:T-n_Diagram}
  \end{figure}

Once the pressure is known at the point of phase transition, for a given 
temperature and number density, the liquid and solid number densities at 
the phase transition can be found from plots of the pressure versus the 
inverse of the number density for both the solid and the liquid, 
as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
find the liquid and solid number densities at the phase transition 
for various temperatures, a Temperature-Density (T-n) phase diagram 
can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
but without the gray hump characteristic of fluids that have 
both attractive and repulsive interatomic potentials.

@@

Classical Density Functional Theory will be used to find the 
Helmholtz free energies of the liquid and the solid for various 
temperatures and number densities, so that phase diagrams can be 
constructed. 
The key idea behind classical Density Functional Theory (cDFT) is that 
free energy can be written as a functional (a function of a function) of 
a number density profile $n(\vec{r})$ which describes the way the number 
density, or atoms per volume, varies spatially on average. 
\begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
This, coupled with the fact that the free energy of a system tends toward 
a minimum at equilibrium means that the equilibrium free energy of the 
system ${f_{eq}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
number density profile $n_{eq}(\vec r)$, can be found simply by varying 
$n(\vec{r})$ through all possible spatial profiles until the free energy 
is minimized~\cite{MoritaDFT}. 
\begin{displaymath}f[n(\vec r)]\rightarrow f_{eq}[n_{eq}(\vec r)]  \mbox{ as $f$ is minimized} \end{displaymath}
\begin{displaymath}{\mbox{where }  n(\vec{r})=n_{eq}(\vec r)  \mbox{ at equilibrium}}\end{displaymath}
The reason the minimization works is because DFT satisfies the conditions 
of the variational principle. 
%the same mathematical principle used to determine 
%that the shortest distance between two points is a straight line by 
%varying the path between the two points until the distance is minimized.

@@

Once an expression for the Helmholtz free energy as a functional of the
number density profile $n(\vec r)$ is formed then, according to classical 
Density Functional Theory, the equilibrium value of the 
Helmholtz energy, and its corresponding equilibrium density profile can be
found by varying the number density profile
through all possible configurations at some fixed temperature
until the free energy is minimized.  
%In this section 
%a functional for the Helmholtz free energy was formed by finding 
%general expressions for the ideal and the excess portions of the free 
%energy, and adding them together.
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
%The functional for the ideal portion of the Helmholtz free energy is given by 
%Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
%free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
%form for the excess Helmholtz free energy depends on the particular 
%fluid being modeled as defined by the  
%interatomic potential energy V(r) between the spheres, which can make
%forming an expression difficult. In this paper, Soft Fundamental Measure
%Theory will be used to form an expression for the excess Helmholtz free 
%energy, as explained in the next section.

\section{Soft Fundamental Measure Theory (SFMT)}
Soft Fundamental Measure Theory (SFMT) is a method for forming an 
expression for the excess Helmholtz free energy of a soft-sphere fluid.
In the next chapter, 
we will adapt an SFMT functional to fit the soft-sphere WCA fluid.
Recall that in order to apply classical Density Functional Theory 
to a system that has a fixed temperature, volume, and number of particles, 
a functional for the Helmholtz free energy must first be formed. 
The ideal portion of the Helmholtz free energy is given by 
Equation~\ref{eq:Fideal-n(r)}, but forming an expression for the excess 
Helmholtz free energy from Equation~\ref{eq:Fexcess-simplified} for the 
particular fluid being modeled can be challenging. 
SFMT is a method for forming an 
expression for the excess Helmholtz free energy
that does not depend on explicitly taking into account the interaction 
that each sphere with every other sphere in the fluid (as required by 
Equation~\ref{eq:Fexcess-simplified}).
%In the next chapter on methods, we will use an SFMT functional that we 
%adapt to fit the soft-sphere WCA fluid.
SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
for soft-spheres, and is an extension of Fundamental Measure Theory (FMT) 
originally developed by 
Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
for the hard-sphere fluid. FMT in turn extended  
Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
developed for homogeneous hard-sphere fluids, 
to inhomogeneous, or non-uniform hard-sphere fluids.

\textcolor{red}{FROM HERE}

@@

%This can easily be shown by putting the interatomic potential between two hard spheres
%given by 
    %V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
%into the equation for the Mayer function
     %f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1

@@

Lifting the restriction to the low density limit, Rosenfeld derived an
expression for the excess Helmholtz free energy by noting that the higher
order terms in the virial expansion can all be written in terms of an 
"exclusion volume" which is given by Equation~\ref{exclusion-volume}, 
and an "overlap volume".
The exclusion volume is the spherical volume about the center of one sphere 
which the center of another sphere can not enter, and is used to describe
terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. The overlap volume is 
the volume of intersection between two spheres that overlap, and is used in conjunction 
with the exclusion volume to describe higher order terms in the expansion~\cite{rosenfeld1988}. 
%
The resulting expression for the excess Helmholtz free energy with an 
updated version of $\Phi_3$ that was later constructed by Tarazona is given by
\begin{align}
    F_{excess}[n(\vec{r})] &= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r} \\
    %\Phi_1 &= -n_{0}ln(1-n_{3})  \\
    %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
    %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}
    \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
    \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
    \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}
where in addition to the weight functions already defined, a tensor weight function is given by
\begin{align}\label{eq:tensorweight-methods}
     \overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
\end{align}
The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
all depend on the spatially varying number density profile $n(\vec r)$.
\begin{align}%\label{eq:numdenprofile}
	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
\end{align}
This result reduces to same result obtained from Scaled Particle Theory 
for a homogeneous hard-sphere fluid.

The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
All of the weight functions can be expressed in terms of 
$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
the relation
\begin{align}\label{w2_w3_relation}
    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
\end{align}
%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
another important relation expressed in terms of $w_{2}(r)$ 
%that will be important 
%to keep in tact when moving to SFMT~\cite{schmidt2000fluid}.
\begin{align}\label{mayerder=w2convolution}
     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
\end{align} 
Soft Fundamental Measure Theory extends FMT to fluids 
with soft potentials by keeping the relationships between the weight 
functions, and the equation for the derivative of the Mayer function 
the same, 
but replacing $w_{3}(r)$ with a weight function appropriate 
for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
This is what we do in forming a functional for the Weeks-Chandler-Anderson 
fluid as detailed in the next chapter on Methods.

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two distinct regions - a "liquid" region at lower densities, and a solid 
region at higher densities. 
A phase transition between a liquid and a solid is accompanied by an 
abrupt change in the number density, and so there is a gap between the 
purely liquid and the purely solid regions. Within the gap, or coexistence
region, is found mixtures of various amounts of liquid and solid each of
which have number densities that are fixed at the values they had at the 
phase transition. 
To construct the phase diagram, the liquid and solid phase transition number densities
are found 
%Maxwell's double tangent construction can be used to find 
%these number densities given plots of 
from the liquid and the solid Helmholtz 
free energies and inverse densities at a given temperature, as
discussed in Section 2.2.

We use classical Density Functional Theory 
to find the Helmholtz free energies of the liquid and the solid. 
A functional for the ideal portion of the
Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, and a 
functional for 
the excess Helmholtz free energy, which depends on the interatomic potential
between particles, will be developed in the next section using SFMT. 
Once a functional is constructed, the Helmholtz free energy can be found - 
but not directly. The equilibrium number density profile must first be 
identified. According to cDFT, the equilibrium number density profile, 
and its corresponding equilibrium Helmholtz free energy can be found 
by trying many different number densities, and identifying the one that 
minimizes the Helmholtz free energy. In this section, we look at how to 
generate various number density profiles that can be used to implement cDFT.

@@

\section{Computing the Liquid and Crystal Helmholtz free energies}

>>>>

%Newer Intro....
%The goal of this project is to generate temperature-density 
%and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
%(WCA) model fluid using classical Density Functional Theory (cDFT). 
%To do this, the Helmholtz free energies of the fluid
%in a fluid state and in a crystalline solid state must be found.
%In general, the Helmholtz free energy, which consists of an ideal part 
%and an excess part, can be written as a functional of the number density 
%profile $n(\vec{r})$ which describes the way the average number 
%density varies spatially. 
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

@@

A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
the density will change abruptly at the phase transition.
The particular structure and density that result are affected 
by the interatomic potential. When repulsive, soft sphere fluids 
(that are not too soft) freeze, they form lattice structures composed of 
Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
This structure is what one would see by filling a box with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

\begin{figure}
    \centering
    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
\end{figure}

A phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure~\ref{fig:P-T_Diagram}. 
%
%Points along the phase boundary that separates the solid and the fluid phases 
%indicate a phase transition where the solid and fluid phases coexist, 
%and show that both the solid and the fluid have the same temperature (necessary 
%for thermal equilibrium)
%and the same pressure (necessary for mechanical 
%equilibrium)at a phase transition, and the same chemical potential 
%(necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some substance 
which exhibits both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will just simply
be a fluid.\textcolor{red}{liquid in this paper.}
The phase diagram indicates the equilibrium state a substance will be in at
different temperatures and pressures.
At a phase tranition, both the fluid and the solid have 
the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
%The temperature , pressure, and chemical potential will be the same for both
%the fluid and the solid at a fluid-solid phase transition.

@@

 \begin{figure}
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
    \caption{Fluid and solid densities at the solid-fluid phase 
    transition are identified using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the fluid
and one for the solid at a fixed temperature. 
%Each curve consists of a collection of data points $(\frac{F}{N}, \frac{1}{n})$ 
The data for the curves 
%and the solid respectively, which 
will be obtained by using cDFT. 
%to find the Helmholtz free energies of the fluid and the solid for various 
%temperatures and number densities. 
Since the pressure of the 
fluid is the same as the pressure of the solid at the phase transition, 
the density of the fluid and the density of the solid at 
freezing can be found from identifying the points at which the 
tangents to each curve have the same slope, as shown by the black line. 
This process is called Maxwell's Double Tangent Construction.

To facilitate finding the pressure at the phase transition computationally, 
it is helpful to take different approach by considering the Gibbs free energy given by G=F+PV. 
Rearranging the terms and dividing by N gives an  expression for the 
Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
per atom  $g$ (which is the chemical potential)  
\begin{equation}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{equation}
\begin{align}f=-P\frac{1}{n}+g\end{align}
This equation has the form of a line
\begin{equation}
    y=mx+y_o
\end{equation}
and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom, or chemical potential given by the y-intercept. 
At the point of phase transition the chemical potential of the fluid is 
the same as the chemical potential of the solid, as is the pressure.
\begin{equation}
    \mu_F=\mu_S\mbox{~~~~and~~~~}P_F=P_S\mbox{~~~~at a phase transition}
\end{equation}

 \begin{figure}
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs 
    when $\mu_F=\mu_S$.}
    \label{fig:GibbsvsP}
  \end{figure}

\begin{figure}
    \centering
    \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
    \caption{The fluid and solid densities at the phase transition are 
    found from a plot of pressure vs inverse density for both the fluid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

The pressure at the phase transition can be obtained from the point of 
intersection that occurs when the chemical potential 
is plotted against the pressure for both the fluid and the 
solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
This plot is constructed from data collected from lines tangent 
along the curves on the Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
fluid and the solid at a given temperature. 
The slope of the tangent line at a given point gives the negative of the 
pressure at that point, and the corresponding y-intercept of the tangent 
line gives the chemical potential at that point. Collecting this data 
(P,~$\mu$) at each point for the fluid and the solid respectively, 
produces the curves in Figure~\ref{fig:GibbsvsP}.

\begin{figure}
    \centering
    \includegraphics[height=7cm]{figs/T-n_Diagram.pdf}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the fluid and the solid during a phase 
    transition. The gray region is where different phases coexist in 
    varying amounts at their phase transition number densities.}
    \label{fig:T-n_Diagram}
  \end{figure}

Once the pressure is known at the point of phase transition, for a given 
temperature and number density, the fluid and solid number densities at 
the phase transition can be found from plots of the pressure versus the 
inverse of the number density for both the solid and the fluid, 
as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
find the fluid and solid number densities at the phase transition 
for various temperatures, a Temperature-Density (T-n) phase diagram 
can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
but without the gray hump characteristic of fluids that have 
attractive interatomic potentials, and which represents the region where
vapor coexists with liquid as evidenced by an interface.

@@

Classical Density Functional Theory will be used to find the 
Helmholtz free energies of the fluid and the solid for various 
temperatures and number densities, so that phase diagrams can be 
constructed. 
The key idea behind classical Density Functional Theory (cDFT) is that 
free energy can be written as a functional (a function of a function) of 
a number density profile $n(\vec{r})$ which describes the way the number 
density, or atoms per volume, varies spatially on average. 
\begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
This, coupled with the fact that the free energy of a system tends toward 
a minimum at equilibrium means that the equilibrium free energy of the 
system ${f_{eq}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
number density profile $n_{eq}(\vec r)$, can be found simply by varying 
$n(\vec{r})$ through all possible spatial profiles until the free energy 
is minimized~\cite{MoritaDFT}. 
\begin{displaymath}f[n(\vec r)]\rightarrow f_{eq}[n_{eq}(\vec r)]  \mbox{ as $f$ is minimized} \end{displaymath}
\begin{displaymath}{\mbox{where }  n(\vec{r})=n_{eq}(\vec r)  \mbox{ at equilibrium}}\end{displaymath}
The reason the minimization works is because DFT satisfies the conditions 
of the variational principle. 
%the same mathematical principle used to determine 
%that the shortest distance between two points is a straight line by 
%varying the path between the two points until the distance is minimized.

@@

Once an expression for the Helmholtz free energy as a functional of the
number density profile $n(\vec r)$ is formed then, according to classical 
Density Functional Theory, the equilibrium value of the 
Helmholtz energy, and its corresponding equilibrium density profile can be
found by varying the number density profile
through all possible configurations at some fixed temperature
until the free energy is minimized.

%In this section 
%a functional for the Helmholtz free energy was formed by finding 
%general expressions for the ideal and the excess portions of the free 
%energy, and adding them together.
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
%The functional for the ideal portion of the Helmholtz free energy is given by 
%Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
%free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
%form for the excess Helmholtz free energy depends on the particular 
%fluid being modeled as defined by the  
%interatomic potential energy V(r) between the spheres, which can make
%forming an expression difficult. In this paper, Soft Fundamental Measure
%Theory will be used to form an expression for the excess Helmholtz free 
%energy, as explained in the next section.

\section{Soft Fundamental Measure Theory (SFMT)}
Soft Fundamental Measure Theory (SFMT) is a method for forming an 
expression for the excess Helmholtz free energy of a soft-sphere fluid.
In the next chapter, 
we will adapt an SFMT functional to fit the soft-sphere WCA fluid.
Recall that in order to apply classical Density Functional Theory 
to a system that has a fixed temperature, volume, and number of particles, 
a functional for the Helmholtz free energy must first be formed. 
The ideal portion of the Helmholtz free energy is given by 
Equation~\ref{eq:Fideal-n(r)}, but forming an expression for the excess 
Helmholtz free energy from Equation~\ref{eq:Fexcess-simplified} for the 
particular fluid being modeled can be challenging. 
SFMT is a method for forming an 
expression for the excess Helmholtz free energy
that does not depend on explicitly taking into account the interaction 
that each sphere with every other sphere in the fluid (as required by 
Equation~\ref{eq:Fexcess-simplified}).
%In the next chapter on methods, we will use an SFMT functional that we 
%adapt to fit the soft-sphere WCA fluid.
SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
for soft-spheres, and is an extension of Fundamental Measure Theory (FMT) 
originally developed by 
Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
for the hard-sphere fluid. 
%FMT in turn extended  
%Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
%developed for homogeneous hard-sphere fluids, 
%to inhomogeneous, or non-uniform hard-sphere fluids.

%\textcolor{red}{FROM HERE}

@@

This can easily be shown by putting the interatomic potential between two hard spheres
$V(r)$
into the equation for the Mayer function
     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
where
    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq R_i+R_j \\ 0  & r> R_i+R_j \end{array}\right.

@@

Lifting the restriction to the low density limit, Rosenfeld derived an
expression for the excess Helmholtz free energy by noting that the higher
order terms in the virial expansion can all be written in terms of an 
"exclusion volume" which is given by Equation~\ref{exclusion-volume}, 
and an "overlap volume".
The exclusion volume is the spherical volume about the center of one sphere 
which the center of another sphere can not enter, and is used to describe
terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. The overlap volume is 
the volume of intersection between two spheres that overlap, and is used in conjunction 
with the exclusion volume to describe higher order terms in the expansion
in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988}. 
%
The resulting expression for the excess Helmholtz free energy with an 
updated version of $\Phi_3$ that was later constructed by Tarazona is given by
\begin{align}
    F_{excess}[n(\vec{r})] &= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r} \\
    %\Phi_1 &= -n_{0}ln(1-n_{3})  \\
    %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
    %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}
    \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
    \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
    \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}
where in addition to the weight functions already defined, a tensor weight function is given by
\begin{align}\label{eq:tensorweight-methods}
     \overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
\end{align}
The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
all depend on the spatially varying number density profile $n(\vec r)$.
\begin{align}%\label{eq:numdenprofile}
	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
\end{align}
This result reduces to same result obtained from Scaled Particle Theory 
for a homogeneous hard-sphere fluid.

The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
All of the weight functions can be expressed in terms of 
$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
the relation
\begin{align}\label{w2_w3_relation}
    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
\end{align}
%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
Taking the derivative of the Mayer function provides 
another important relation expressed in terms of $w_{2}(r)$ 
%that will be important 
%to keep in tact when moving to SFMT~\cite{schmidt2000fluid}.
\begin{align}\label{mayerder=w2convolution}
     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
\end{align} 
Soft Fundamental Measure Theory extends FMT to fluids 
with soft potentials by keeping the relationships between the weight 
functions, and the equation for the derivative of the Mayer function 
the same, 
but replacing $w_{3}(r)$ with a weight function appropriate 
for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
This is what we do in forming a functional for the Weeks-Chandler-Anderson 
fluid as detailed in the next chapter on Methods.

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two distinct regions - a fluid region at lower densities, and a solid 
region at higher densities. 
A phase transition between a fluid and a solid is accompanied by an 
abrupt change in the number density, and so there is a gap between the 
purely fluid and the purely solid regions. Within the gap, or coexistence
region, is found mixtures of various amounts of fluid and solid each of
which have number densities that are fixed at the values they had at the 
phase transition. 
To construct the phase diagram, the fluid and solid phase transition 
number densities are found 
%Maxwell's double tangent construction can be used to find 
%these number densities given plots of 
from the fluid and the solid Helmholtz 
free energies and inverse densities at a given temperature, as
discussed in Section 2.2.

We use classical Density Functional Theory 
to find the Helmholtz free energies of the fluid and the solid. 
A functional for the ideal portion of the
Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, and a 
functional for 
the excess Helmholtz free energy, which depends on the interatomic potential
between particles, will be developed in the next section using SFMT. 
Once a functional is constructed, the Helmholtz free energy can be found - 
but not directly. The equilibrium number density profile must first be 
identified. According to cDFT, the equilibrium number density profile, 
and its corresponding equilibrium Helmholtz free energy can be found 
by trying many different number densities, and identifying the one that 
minimizes the Helmholtz free energy. In this section, we look at how to 
generate various number density profiles that can be used to implement cDFT.

@@

\section{Computing the Fluid and Crystal Helmholtz free energies}

The fluid ideal Helmholtz free energy is computed analytically 
using Equation~\ref{eq:Fideal} 
for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
The crystal ideal Helmholtz free energy is computed computationally 
from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
\begin{align}
  n(|\vec{r}-\vec{R_j}|)=\sum_j{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_j}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}
\end{align}
where $\vec R_j$ is the position vector of the jth lattice site.
The fluid excess Helmholtz free energy is computed analytically  
using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
with a uniform number density profile, n that does not depend on $\vec r$.
